---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
```

### Load data

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data

BRFSS conducts both landline telephone and cellular telephone based surveys. In conducting the BRFSS landline telephone survey, interviewers collect data from a randomly selected adult in a household. In conducting the cellular telephone version of the BRFSS questionnaire, interviewers collect data from an adult who participates by using a cellular telephone and resides in a private residence or college housing. The telephone numbers dialed are generated using random number generation.

Health characteristics estimated from the BRFSS pertain to the non-institutionalized adult population, aged 18 years or older, who reside in the US.

Using this methodology helps control selection bias. However the survey is somewhat lengthy (often around 45 minutes to complete), and additionally pertains to potentaially sensitive topics which may introduce non-response bias.

This is an observational study over a limited point in time.

* * *

## Part 2: Research questions
The following research questions pertain to sleep amount.

**Research quesion 1:**
Is there a relationship between time slept and alcohol consumption?

**Research quesion 2:**
Is there a relationship between time slept and number of children in the household?

**Research quesion 3:**
Is there a relationship between time slept and the type of payment for work?

* * *

## Part 3: Exploratory data analysis

Create dataset with relavent variables
```{r}
brfss_clean <- brfss2013 %>%
        select(sleptim1, avedrnk2, children, scntlpad)
```
**Research quesion 1:**

```{r}
brfss_sl_alc <- brfss_clean %>%
        select(sleptim1, avedrnk2) %>%
        drop_na()

summary(brfss_sl_alc)

p_alc <- ggplot(data = brfss_sl_alc, aes(x = sleptim1, y = avedrnk2))

p_alc + geom_point() + geom_smooth()
```



**Research quesion 2:**

```{r}
#prepare dataset, remove obvious outliers and na
brfss_sl_chl <- brfss2013 %>%
        select(sleptim1, children, X_ageg5yr) %>%
        drop_na() %>%
        filter(children < 20 & sleptim1 < 24)

#look at the data with tables and numerical summary
head(brfss_sl_chl)

summary(brfss_sl_chl)

brfss_sl_chl %>%
  group_by(X_ageg5yr) %>%
  summarise(mean = mean(children), median = median(children))

#look at the data visually
p_chl <- ggplot(data = brfss_sl_chl, aes(x = children, y = sleptim1))

p_chl + geom_point()

p_chl + geom_jitter() + coord_flip()

ggplot(data = brfss_sl_chl, aes(x = sleptim1, y = children)) +
  geom_point()

ag_ch <- ggplot(data = brfss_sl_chl, aes(x = X_ageg5yr, y = children))
  
ag_ch + geom_boxplot() + coord_flip()

a_c <- brfss_sl_chl %>%
  group_by(X_ageg5yr) %>%
  mutate(outlier_high = children > quantile(children, .75) + 1.50*IQR(children), outlier_low = children < quantile(children, .25) - 1.50*IQR(children)) %>%
  ungroup()

a_c <- a_c %>%
  group_by(X_ageg5yr) %>%
  mutate(outlier = children > median(children) + IQR(children) * 1.5) %>%
  ungroup()

a_c <- a_c %>%
  mutate(outlier_color = case_when(outlier_high ~ "red", outlier_low ~ "steelblue"))

head(a_c)


ggplot(data = a_c, aes(x = X_ageg5yr, y = children)) +
         geom_boxplot(outlier.shape = NA) + coord_flip() + geom_jitter(color = a_c$outlier_color, width = .2)

ggplot(data = a_c, aes(x = X_ageg5yr, y = children)) +
         geom_boxplot(outlier.shape = NA) + coord_flip() +
          geom_jitter(data = filter(a_c, outlier ==T), width = .2)

```



**Research quesion 3:**

```{r}
brfss_sl_pay <- brfss_clean %>%
        select(sleptim1, scntlpad) %>%
        drop_na()

brfss_sl_pay %>%
  group_by(scntlpad) %>%
  summarise(median(sleptim1), mean(sleptim1), IQR(sleptim1))

p_pay <- ggplot(data = brfss_sl_pay, aes(x = scntlpad, y = sleptim1))

p_pay + geom_boxplot()
```

